Placeholder for article 3

### 2025-02-26 analysis/paper/paper.qmd
Macbook morning work session. Changes to tables in main paper and supplementary.

# Tables with class by raw material type (n and %)
##
general_table <- phase_dataset |> 
    dplyr::mutate(RMTYPE = case_when(
    RMTYPE == "TYPE 1" | RMTYPE == "TYPE 2" | RMTYPE == "TYPE 3" | RMTYPE == "TYPE 4" | RMTYPE == "TYPE 5" ~ "Local",
    RMTYPE == "TYPE 6" ~ "T6",
    RMTYPE == "TYPE 7" ~ "T7",
    RMTYPE == "TYPE 8" ~ "T8",
    RMTYPE == "TYPE 9" ~ "T9",
    RMTYPE == "TYPE 10" ~ "T10",
    TRUE ~ RMTYPE)) |> 
  dplyr::mutate(CLASS = case_when(
    CLASS == "BlankFrag" ~ "Blank fragment",
    CLASS == "BurinSpall" ~ "Burin spall",
    CLASS == "CoreFrag" ~ "Core fragment",
    CLASS == "CorePreparProd" ~ "Core prep product",
    CLASS == "RetouchedPiece" ~ "Retouched",
    CLASS == "RetouchedPieceFrag" ~ "Retouched fragment",
    TRUE ~ CLASS)) |> 
  dplyr::mutate(RETOUCHEDPIECETYPOLOGY = case_when(
    RETOUCHEDPIECETYPOLOGY == "Retouched blade" | RETOUCHEDPIECETYPOLOGY == "Retouched bladelet" | RETOUCHEDPIECETYPOLOGY == "Retouched bladelet fragment" | RETOUCHEDPIECETYPOLOGY == "Retouched flake" | RETOUCHEDPIECETYPOLOGY == "Retouched flake fragment" ~ "Retouched blank",
    RETOUCHEDPIECETYPOLOGY == "Splintered piece fragment" ~ "Splintered piece",
    TRUE ~ RETOUCHEDPIECETYPOLOGY))

@@

# Solutrean table
## Table preparation
solutrean_retouch <- general_table_solutrean |>
  filter(CLASS %in% c("Retouched", "Retouched fragment")) |>
  filter(RMTYPE != "INDET") |> 
  filter(RETOUCHEDPIECETYPOLOGY != "") |> 
  tabyl(RETOUCHEDPIECETYPOLOGY, RMTYPE) |>
  adorn_totals(c("row", "col")) |>
  mutate(across(where(is.numeric), ~ ifelse(. == 0, " ", .))) |>
  select(RETOUCHEDPIECETYPOLOGY, Local, T6, T7, T8, T11, Total) |> # Ordering values
  rename(Typology = RETOUCHEDPIECETYPOLOGY)

@@

>>>>

# Tables with class by raw material type (n and %)
##
general_table <- phase_dataset |> 
    dplyr::mutate(RMTYPE = case_when(
    RMTYPE == "TYPE 1" | RMTYPE == "TYPE 2" | RMTYPE == "TYPE 3" | RMTYPE == "TYPE 4" | RMTYPE == "TYPE 5" ~ "Local",
    RMTYPE == "TYPE 6" ~ "T6",
    RMTYPE == "TYPE 7" ~ "T7",
    RMTYPE == "TYPE 8" ~ "T8",
    RMTYPE == "TYPE 9" ~ "T9",
    RMTYPE == "TYPE 10" ~ "T10",
    TRUE ~ RMTYPE)) |> 
  dplyr::mutate(CLASS = case_when(
    CLASS == "BlankFrag" ~ "Blank fragment",
    CLASS == "BurinSpall" ~ "Burin spall",
    CLASS == "CoreFrag" ~ "Core fragment",
    CLASS == "CorePreparProd" ~ "Core prep product",
    CLASS == "RetouchedPiece" ~ "Retouched",
    CLASS == "RetouchedPieceFrag" ~ "Retouched fragment",
    TRUE ~ CLASS)) |> 
  dplyr::mutate(RETOUCHEDPIECETYPOLOGY = case_when(
    RETOUCHEDPIECETYPOLOGY == "Retouched blade" | RETOUCHEDPIECETYPOLOGY == "Retouched bladelet" | RETOUCHEDPIECETYPOLOGY == "Retouched bladelet fragment" | RETOUCHEDPIECETYPOLOGY == "Retouched flake" | RETOUCHEDPIECETYPOLOGY == "Retouched flake fragment" | RETOUCHEDPIECETYPOLOGY == "Retouched burin spall" ~ "Retouched blank",
    RETOUCHEDPIECETYPOLOGY == "Splintered piece fragment" ~ "Splintered piece",
    RETOUCHEDPIECETYPOLOGY == "Backed bladelet" | RETOUCHEDPIECETYPOLOGY == "Backed flake" ~ "Backed tool",
    RETOUCHEDPIECETYPOLOGY == "Retouched burin" ~ "Burin",
    TRUE ~ RETOUCHEDPIECETYPOLOGY))

@@

# Solutrean table
## Table preparation
solutrean_retouch <- general_table_solutrean |>
  filter(CLASS %in% c("Retouched", "Retouched fragment")) |>
  dplyr::mutate(RETOUCHEDPIECETYPOLOGY = case_when(
    RETOUCHEDPIECETYPOLOGY == "Bifacial fragment" ~ "Bifacial blank",
    RETOUCHEDPIECETYPOLOGY == "Shouldered point fragment" ~ "Shouldered point",
    TRUE ~ RETOUCHEDPIECETYPOLOGY))
  filter(RMTYPE != "INDET") |> 
  filter(RETOUCHEDPIECETYPOLOGY != "") |> 
  tabyl(RETOUCHEDPIECETYPOLOGY, RMTYPE) |>
  adorn_totals(c("row", "col")) |>
  mutate(across(where(is.numeric), ~ ifelse(. == 0, " ", .))) |>
  select(RETOUCHEDPIECETYPOLOGY, Local, T6, T7, T8, T11, Total) |> # Ordering values
  rename(Typology = RETOUCHEDPIECETYPOLOGY)

@@

```{r}
#| label: stats-wilcox

# Terrace wilcox statistics
## DB prep
wilcox_t <- phase_dataset |> 
  dplyr::mutate(Prov = case_when(
    RMTYPE == "TYPE 1" | RMTYPE == "TYPE 2" | RMTYPE == "TYPE 3" | RMTYPE == "TYPE 4" | RMTYPE == "TYPE 5" ~ "Local",
    RMTYPE == "TYPE 6" | RMTYPE == "TYPE 7" | RMTYPE == "TYPE 8" | RMTYPE == "TYPE 9" ~ "Non-local",
    TRUE ~ RMTYPE)) |> 
  filter(Prov %in% c("Local", "Non-local")) |> 
  select(ID, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT, Prov, Level)

# Proto-Solutrean blanks statistical test
## NL cores were not represented enough (n) for statistics
wt_blank_proto <- wilcox_t |> 
  filter(Level %in% c("4E", "4E/5")) |> 
  filter(CLASS == "Blank")

## Blanks weight (bw)
dplyr::group_by(wt_blank_proto, Prov) |> 
  summarise(count = n(), median = median(WEIGHT, na.rm = TRUE))
res_bw_proto <- wilcox.test(WEIGHT ~ Prov, data = wt_blank_proto, exact = FALSE)
## Blanks length (bl)
dplyr::group_by(wt_blank_proto, Prov) |> 
  summarise(count = n(), median = median(LENGTH, na.rm = TRUE))
res_bl_proto <- wilcox.test(LENGTH ~ Prov, data = wt_blank_proto, exact = FALSE)
## Blanks width (bw)
dplyr::group_by(wt_blank_proto, Prov) |> 
  summarise(count = n(), median = median(MAXWIDTH, na.rm = TRUE))
res_bwi_proto <- wilcox.test(MAXWIDTH ~ Prov, data = wt_blank_proto, exact = FALSE)

# Gravettian blanks statistical test
wt_blank_grav <- wilcox_t |> 
  filter(Level %in% c("6", "7")) |> 
  filter(CLASS == "Blank")

## Blanks weight
dplyr::group_by(wt_blank_grav, Prov) |> 
  summarise(count = n(), median = median(WEIGHT, na.rm = TRUE))
res_bw_grav <- wilcox.test(WEIGHT ~ Prov, data = wt_blank_grav, exact = FALSE)
## Blanks length
dplyr::group_by(wt_blank_grav, Prov) |> 
  summarise(count = n(), median = median(LENGTH, na.rm = TRUE))
res_bl_grav <- wilcox.test(LENGTH ~ Prov, data = wt_blank_grav, exact = FALSE)
## Blanks width
dplyr::group_by(wt_blank_grav, Prov) |> 
  summarise(count = n(), median = median(MAXWIDTH, na.rm = TRUE))
res_bwi_grav <- wilcox.test(MAXWIDTH ~ Prov, data = wt_blank_grav, exact = FALSE)

## Gravettian cores statistical test
wt_core_grav <- wilcox_t |> 
  filter(Level %in% c("6", "7")) |> 
  filter(CLASS == "Core")

## Cores weight
dplyr::group_by(wt_core_grav, Prov) |> 
  summarise(count = n(), median = median(WEIGHT, na.rm = TRUE))
res_cw_grav <- wilcox.test(WEIGHT ~ Prov, data = wt_core_grav, exact = FALSE)
## Cores length
dplyr::group_by(wt_core_grav, Prov) |> 
  summarise(count = n(), median = median(LENGTH, na.rm = TRUE))
res_cl_grav <- wilcox.test(LENGTH ~ Prov, data = wt_core_grav, exact = FALSE)
## Cores width
dplyr::group_by(wt_core_grav, Prov) |> 
  summarise(count = n(), median = median(MAXWIDTH, na.rm = TRUE))
res_cwi_grav <- wilcox.test(MAXWIDTH ~ Prov, data = wt_core_grav, exact = FALSE)

# Shelter wilcox statistics
## DB prep
wilcox_t_shelter <- solutrean_dataset |> 
  dplyr::mutate(Prov = case_when(
    RMTYPE == "TYPE 1" | RMTYPE == "TYPE 2" | RMTYPE == "TYPE 3" | RMTYPE == "TYPE 4" | RMTYPE == "TYPE 5" ~ "Local",
    RMTYPE == "TYPE 6" | RMTYPE == "TYPE 7" | RMTYPE == "TYPE 8" | RMTYPE == "TYPE 9" | RMTYPE == "TYPE 11" ~ "Non-local",
    TRUE ~ RMTYPE)) |> 
   mutate(CLASS = case_when(
    CLASS == "Flake" | CLASS == "ElongBlank" | CLASS == "Thinning flake" ~ "Blank",
    TRUE ~ CLASS)) |>
  filter(Prov %in% c("Local", "Non-local")) |> 
  select(ID, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT, Prov, LEVEL)

# Solutrean blanks statistical test
wt_blank_s <- wilcox_t_shelter |>
  filter(CLASS == "Blank")

## Blanks weight
dplyr::group_by(wt_blank_s, Prov) |> 
  summarise(count = n(), median = median(WEIGHT, na.rm = TRUE))
res_bw_s <- wilcox.test(WEIGHT ~ Prov, data = wt_blank_s, exact = FALSE)
## Blanks length
dplyr::group_by(wt_blank_s, Prov) |> 
  summarise(count = n(), median = median(LENGTH, na.rm = TRUE))
res_bl_s <- wilcox.test(LENGTH ~ Prov, data = wt_blank_s, exact = FALSE)
## Blanks width
dplyr::group_by(wt_blank_s, Prov) |> 
  summarise(count = n(), median = median(MAXWIDTH, na.rm = TRUE))
res_bwi_s <- wilcox.test(MAXWIDTH ~ Prov, data = wt_blank_s, exact = FALSE)

wt_core_s <- wilcox_t_shelter |>
  filter(CLASS == "Core")

## Cores weight
dplyr::group_by(wt_core_s, Prov) |> 
  summarise(count = n(), median = median(WEIGHT, na.rm = TRUE))
res_cw_s <- wilcox.test(WEIGHT ~ Prov, data = wt_core_s, exact = FALSE)
## Cores length
dplyr::group_by(wt_core_s, Prov) |> 
  summarise(count = n(), median = median(LENGTH, na.rm = TRUE))
res_cl_s <- wilcox.test(LENGTH ~ Prov, data = wt_core_s, exact = FALSE)
## Cores width
dplyr::group_by(wt_core_s, Prov) |> 
  summarise(count = n(), median = median(MAXWIDTH, na.rm = TRUE))
res_cwi_s <- wilcox.test(MAXWIDTH ~ Prov, data = wt_core_s, exact = FALSE)

p.value_blanks <- tibble(
  Technocomplex = c("Gravettian", " "," ", " ", "Proto-Solutrean", " ", " ", " ", "Solutrean", " ", " ", " "),
  Variable = c("Weight", "Length", "Width", "(n)","Weight", "Length", "Width","(n)", "Weight", "Length", "Width", "(n)"),
  L_median = c("2.8", "25.8", "19.1", "82", "2.2", "23.4", "16.8", "154", "1.7", "25.1", "16.6", "342"),
  NL_median = c("1.9", "21.2", "17.8", "69", "1.7", "19.4", "16.6", "49", "1.5", "21.9", "18.8", "86"),
  P_Value = round(c(res_bw_grav$p.value, res_bl_grav$p.value, res_bwi_grav$p.value, 0, res_bw_proto$p.value, res_bl_proto$p.value, res_bwi_proto$p.value, 0, res_bw_s$p.value, res_bl_s$p.value, res_bwi_s$p.value, 0), 5)) |>   mutate(Significance = case_when(
      P_Value <= 0.001 ~ "***",
      P_Value <= 0.01  ~ "**",
      P_Value <= 0.05  ~ "*",
      TRUE             ~ ""),
    P_value = paste0(format(P_Value, nsmall = 4), Significance)  # Keep decimal places
  ) |> 
  select(-P_Value, -Significance) |> 
  dplyr::mutate(P_value = case_when(
    P_value == "0.00000***" ~ "-",
    TRUE ~ P_value))

p.value_cores <- tibble(
  Technocomplex = c("Gravettian", "", "", "", "Solutrean", "", "", ""),
  Variable = c("Weight", "Length", "Width", "(n)", "Weight", "Length", "Width", "(n)"),
  L_median = c("20", "29.6", "27.6", "14", "16.5", "28.3", "30.7", "59"),
  NL_median = c("11.8", "21.9", "29.7", "10", "11.7", "25.6", "26.0", "7"),
  P_Value = round(c(res_cw_grav$p.value, res_cl_grav$p.value, res_cwi_grav$p.value, 0, res_cw_s$p.value, res_cl_s$p.value, res_cwi_s$p.value, 0), 3)) |> 
  mutate(Significance = case_when(
      P_Value <= 0.001 ~ "***",
      P_Value <= 0.01  ~ "**",
      P_Value <= 0.05  ~ "*",
      TRUE             ~ ""),
    P_value = paste0(format(P_Value, nsmall = 3), Significance)  # Keep decimal places
  ) |> 
  select(-P_Value, -Significance) |> 
  dplyr::mutate(P_value = case_when(
    P_value == "0.000***" ~ "-",
    TRUE ~ P_value))
